---
title: 'Practical Machine Learning: Course Project'
author: "Steve Phillips"
date: "Sunday, December 27, 2015"
output:
  html_document:
    pandoc_args: [
      "+RTS", "-K64m",
      "-RTS"
    ]
---

Executive Summary:
==================
The intent of this project was to create a prediction model that could accurately classify the manner in which participants performed a barbell lift using accelerometer data into one of 5 categories.  Using 19,622 records of training data I selected the 59 fields that contained no NA or missing values and excluded the index column.  I built a Random Forest model and performed K Fold cross validation (K = 10) using 75% of the data each time as the training set and 25% as the test set.  The out of sample accuracy ranged between 99.80% and 99.98% with a mean accuracy of 99.88% accuracy.


Results:
=========


The below table summarizes the accuracy for 10 runs of K fold cross validation
```{r MainChunk, cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
library(caret)
library(ggplot2)
library(randomForest)


setwd("F:/Users/SJP/My Documents/R Working Directory")

dataTrain <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",stringsAsFactors = TRUE)
#New line
#train_classes <- unlist(sapply(dataTrain, class))

dataTest <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",stringsAsFactors = TRUE)
inscope <- read.csv("workout_inscope.csv")

liteTrain <- dataTrain[,inscope$InScope]
liteFinalTest <- dataTest[,inscope$InScope]
# new line
names(liteFinalTest)[59] <- "classe"
liteFinalTest$classe <- as.factor(rep("A",20))





# convert time stamp to date and time
liteTrain$cvtd_timestamp <- as.POSIXct(liteTrain$cvtd_timestamp, format = "%d/%m/%Y %H:%M")
liteFinalTest$cvtd_timestamp <- as.POSIXct(liteFinalTest$cvtd_timestamp, format = "%d/%m/%Y %H:%M")

liteFinalTest$magnet_dumbbell_z <- as.numeric(liteFinalTest$magnet_dumbbell_z)
liteFinalTest$magnet_forearm_y <- as.numeric(liteFinalTest$magnet_forearm_y)
liteFinalTest$magnet_forearm_z <- as.numeric(liteFinalTest$magnet_forearm_z)


# make levels of factor variables the same between training and test set
levels(liteFinalTest$new_window) <- levels(liteTrain$new_window)
levels(liteFinalTest$classe) <- levels(liteTrain$classe)


#---------------
# Test data types to make sure that they all match
# a1 <- unlist(sapply(liteTrain, class))
# a2 <- unlist(sapply(liteFinalTest, class))
# a1 == a2
# data.frame(a1,a2,a1==a2)


#---------------






#run K fold cross validation, K = 10

K = 10


#initialize data frame for Accuracy results

df <- data.frame(Accuracy = rep(0.0001,K))

for (i in 1:K) {

  # subset Training data into Training and test set
  
  inTrain <- createDataPartition(y=liteTrain$classe,p=.75,list=FALSE)
  training <- liteTrain[inTrain,]
  testing <- liteTrain[-inTrain,]
  
  
  fit <- randomForest(classe ~ ., data = training)

  
  
  predictions <- predict(fit,testing)
  results <- confusionMatrix(predictions,testing$classe)
  raccuracy <- results$overall[1]
  df$Accuracy[i] <- raccuracy

}

#print("Summary of K fold cross validation accuracy")
print(summary(df$Accuracy))

#print("Output of last random forest model fit")
#print(fit)

#-----------
# predict on final data


fpredictions <- predict(fit,liteFinalTest)

#print("Predictions for 20 test cases")
#print(fpredictions)


```


Predictions for 20 test cases.  Accuracy when submitted was 20 out of 20 correct, for 100% accuracy.  This was within range of predicted accuracy based on K Fold cross validation.

```{r}
print(fpredictions)
```
